---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Appendix 2: Simulations of Model Performance Bias and Variance by Resampling Techniques



## General Setup

Load libraries
```{r}
#| message: false

library(tidyverse)
library(tidymodels)

theme_set(theme_classic())
```

## DGP
Function to generate `n_obs` of simulated observations

* DGP is linear on all `x` + normal error with sd = `irr_err`
* y is simple sum such that all coefficients = 1
* features are correlated based on `sigma`
```{r}
simulate_DGP <- function (n_obs, n_features, irr_err, mu, sigma){

  x <- MASS::mvrnorm(n_obs, mu = mu, Sigma = sigma) |> 
    magrittr::set_colnames(str_c("x", 1:n_features)) |> 
    as_tibble()
  
  x |> 
     mutate(y = rowSums(t(t(x)*b)) + rnorm(n_obs, 
                                        mean = 0, 
                                        sd = irr_err))
}
```


## Simulation settings
```{r}
n_sims <- 1000  # number of simulations
n_obs <- 300 # number of observations

n_features <- 20
mu <- rep(0, n_features)
sigma <- matrix(.3, nrow = n_features, ncol = n_features)
diag(sigma) <- 1
irr_err <- 10
b <- rep(0.5, n_features) # no b0 so set to 0
  
set.seed(123456)
# first call so that we can set up recipe
df <- simulate_DGP(n_obs, n_features, irr_err, mu, sigma) 
rec <- recipe(y ~ ., data = df)

rmse_combined = NULL  # to store RMSE for all methods
```


## Get simulation dfs

```{r}
dfs <- rep(n_obs, n_sims) |>
  map(\(n_obs) simulate_DGP(n_obs, n_features, irr_err, mu, sigma))
```


## What is the TRUE model performance

The irreducible error is set to 10 but our model will have some reducible error too so the true performance of our model will be worse than 10

* The DGP is linear 
* but we only have `r n_obs` observations 
* and we have 20 features

Lets fit many models of n = `r n_obs` (the size of our final model) and assess performance in really big samples of new data (ah the luxury of simulated data!)
```{r}

# models based on full n for each simulation run
models <- dfs |> 
  map(\(df) linear_reg() |> fit(y ~ ., data = df))

# big samples of held out data for high precision assessment of models
# 1 for each model
outs <- rep(10000, n_sims) |> 
  map(\(n_obs) simulate_DGP(n_obs, n_features, irr_err, mu, sigma))

# list of predictions for out from each model
preds <- map2(models, outs, \(model, out) predict(model, out))

# get mean rmse in big held out data set for 1000 full n models
# should be very precise
rmse_true <- map2_dbl(outs, preds, \(out, pred)  rmse_vec(out$y,
                                                      pred$.pred)) |> 
  mean()

message("True RMSE = ", rmse_true)
```

Parallel processing for resampling methods with `fit_resamples()`
```{r}
cl <- parallel::makePSOCKcluster(parallel::detectCores(logical = FALSE))
doParallel::registerDoParallel(cl)
```

## Validation set

### 80/20 split


Here we simulate repeated use of the validation set approach to assess our model performance
```{r}
rmse_combined <- dfs |> 
  map(\(df) validation_split(df, prop = .80)) |>   # validation set split
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarise = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "val_set_80") |>       # label results in df
  bind_rows(rmse_combined)
```

## 50/50 split

Here we simulate repeated use of the validation set approach to assess our model performance
```{r}
rmse_combined <- dfs |> 
  map(\(df) validation_split(df, prop = .50)) |>   # validation set split
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "val_set_50") |>       # label results in df
  bind_rows(rmse_combined)
```

## K-fold

### Simple 5-fold
```{r}
rmse_combined <- dfs |> 
  map(\(df) vfold_cv(df, v = 5)) |>   # 5-fold
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "5-fold") |>       # label results in df
  bind_rows(rmse_combined)
```

### Simple 10-fold
```{r}
rmse_combined <- dfs |> 
  map(\(df) vfold_cv(df, v = 10)) |>   # 10-fold
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "10-fold") |>       # label results in df
  bind_rows(rmse_combined)
```


### 3x 10-Fold
```{r}
rmse_combined <- dfs |> 
  map(\(df) vfold_cv(df, v = 10, repeats = 3)) |>   # 3x10-fold
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "3x10-fold") |> # label results in df
  bind_rows(rmse_combined)
```



## Bootstrap Resampling

### 10 resamples

```{r}
rmse_combined <- dfs |> 
  map(\(df) bootstraps(df, times = 10)) |>   # 10 boots
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "boot_10") |>       # label results in df
  bind_rows(rmse_combined)
```

### 100 resamples
```{r}
rmse_combined <- dfs |> 
  map(\(df) bootstraps(df, times = 100)) |>   # 100 boots
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "boot_100") |>       # label results in df
  bind_rows(rmse_combined)
```

### 1000 resamples
```{r}
rmse_combined <- dfs |> 
  map(\(df) bootstraps(df, times = 1000)) |>   # 1000 boots
  map(\(split) linear_reg() |> fit_resamples(resamples = split, 
                                              preprocessor = rec,
                                              metrics = metric_set(rmse))) |> 
  map(\(fits) collect_metrics(fits, summarize = TRUE)) |> 
  list_rbind() |> 
  mutate(method = "boot_1000") |>       # label results in df
  bind_rows(rmse_combined)
```


## Summarize
```{r a102-summary}
rmse_combined |> 
  group_by(method) |> 
  summarize(rmse_mean = mean(mean),
            rmse_sd = sd(mean),
            n = n())
```


##  Plot sampling distributions
```{r a102-plot}
rmse_combined |> 
  ggplot(aes(x = mean, color = method)) + 
  geom_density() +
  geom_vline(aes(xintercept = mean(rmse_true)),
            color = "blue", linetype = "dashed", linewidth = 1)
```
