---
editor_options: 
  chunk_output_type: console
---

# Principal Components Analysis 

##  Overview
To start, the introductory sections about PCA in [Wikipedia](https://en.wikipedia.org/wiki/Principal_component_analysis) provide a nice orientation and summary of the technique.

In short, PCA 



## A Two Dimensional Example

This example is based loosely on tutorial and demonstration data developed by [Lindsay Smith](https://www.iro.umontreal.ca/~pift6080/H09/documents/papers/pca_tutorial.pdf)


### The original variables

Let's start with a toy dataset for two variables, `x1` and `x2` and a sample size of 10.  We will work in two dimensions to make the example easier to visualize but the generalization to n dimensions (where n = number of variables) is not difficult.

```{r}
library(tidyverse)
d <- tibble(x1 = c(2.5, 0.5, 2.2, 1.9, 3.1, 2.3, 2.0, 1.0, 1.5, 1.1),
            x2 = c(2.4, 0.7, 2.9, 2.2, 3.0, 2.7, 1.6, 1.1, 1.6, 0.9))
d |> print()
```

Here is a scatterplot of the data.   I've added the x and y axes of the coordinate system to make those explicit since the PCA will be rotating this coordinate system to achieve its goals.

```{r}
theme_set(theme_classic() + theme(axis.line=element_blank()))
d |> ggplot(aes(x = x1, y = x2)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "blue") +
  geom_vline(xintercept = 0, color = "blue") +
  xlim(-1,4) +
  ylim(-1,4) +
  theme(axis.text.x = element_text(vjust=-.2))
```



