---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Logistic Regression

```{r}
#| echo: FALSE
#| message: FALSE

library(tidyverse) 
library(tidymodels)
options(conflicts.policy = "depends.ok")
# library(baguette)
theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

Generate a dataset where x is positively related to event 1.  As x increases p(1) increases
```{r}
set.seed(123)
n <- 500 
beta <- 0.5
intercept <- 0 

data <- tibble(x = rnorm(n, mean = 0, sd = 1)) |> 
  mutate(eta = intercept + beta * x,
         p = 1 / (1 + exp(-eta)),
         y = rbinom(n, 1, p)) |> 
  select(-eta, -p)
```

### tidymodel
set the event (1) as the first level of the factor (default for tidymodels)

```{r}
data <- data |> 
  mutate(y_event_first = factor(y, levels = c(1, 0)))
```

Fit the model and look at coefficients.   Coefficent of x is negative.   That seems counter-intutiive to what we would expect if the coefficient for x is telling us the relationship between x and p(event)
```{r}
fit_tm <-
  logistic_reg() |> 
  set_engine("glm") |> 
  fit(y_event_first ~ x, data)

fit_tm |> tidy()
```

```{r}
sens_vec(data$y_event_first, predict(fit_tm, new_data = data)$.pred_class)
```



```{r}
glm (y ~ x, data = data, family = "binomial") |> 
  broom::tidy()
```




